---
title: "data-analysis-replication"
output: html_document
date: "2024-03-30"
---

Necessary Packages
```{r}
library(tidyverse)
library(dplyr)
library(mosaic)
library(radiant)
library(sciplot)
library(infer)
library(tidylog)
library(data.table)
library(readxl)
library(gt)
library(gtsummary)
```

Loading in data
```{r}
f <- "data/Dunham et al. (2020) dataset.xlsx"
d <- read_excel(f,sheet = 1, col_names = TRUE)
```

Substrate diameter 
```{r}
d <- d |> mutate(relBrnchDiamCategories = case_when(
  relBrnchDiamNsTl > 1 ~ "Large",
  relBrnchDiamNsTl >= 0.5 & relBrnchDiamNsTl <= 1 ~ "Medium",
  relBrnchDiamNsTl < 0.5 ~ "Small",
  TRUE ~ "NA"
  ))
```


Substrate orientation
```{r}
d$Sin_SbstOrient_Numeric <- as.numeric(d$Sin_SbstOrient)
d <- d |> drop_na("Sin_SbstOrient_Numeric")
d$Angle_SbstOrient <- (asin(d$Sin_SbstOrient_Numeric))*(180/pi)

d <- d |> mutate(Angle_SbstOrientCategories = case_when(
  Angle_SbstOrient >= -30 & Angle_SbstOrient <= 30 ~ "Horizontal",
  Angle_SbstOrient < -30 ~ "Decline",
  Angle_SbstOrient > 30 ~ "Incline",
  TRUE ~ "NA"
  ))
```

Binomial nomenclature
```{r}
d <- d |> mutate(Species = case_when(
  Genus == "Cebuella" ~ "pygmaea",
  Genus == "Saguinus" ~ "tripartitus",
  TRUE ~ "NA"
  ))
d <- d |> mutate(Binomial = paste(Genus, Species, sep = " "))
```


We used linear mixed models to examine the effects of substrate
diameter and substrate orientation on continuous kinematic
variables, including relative speed, mean DF, mean NSL, relative
FLLD, relative HLLD and percent aerial phase. Relative speed was
entered as a covariate in statistical models for mean DF, mean NSL,
relative FLLD, relative HLLD and percent aerial phase in order to
control for speed-related effects on these kinematic variables.
Individual primate was nested within video clip as a random factor
(intercept) in each model (i.e. some video clips contained multiple
individuals)
```{r}


```

